The guts of the dispRity package

### 2023-12-06 inst/gitbook/03_specific-tutorials.Rmd
Up manual and vignettes
updated book

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
1 | `disalignment` | The rejection of the centroid of a matrix from the major axis of another (typically an `"as.covar"` metric) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
1 | `projections.between` | `projections` metric applied between groups | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
1 | `roundness` | The integral of the ranked scaled eigenvalues of a variance-covariance matrix | `dispRity`
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

\begin{equation}
    ellipse.volume = \frac{\pi^{d/2}}{\Gamma(\frac{d}{2}+1)}\displaystyle\prod_{i=1}^{d} (\lambda_{i}^{0.5})
\end{equation}

@@

#### Volumes and surface metrics
The functions `ellipse.volume`, `convhull.surface`, `convhull.volume` and `n.ball.volume`  all measure the surface or the volume of the ordinated space occupied:

@@

```{r}
## Calculating the ellipsoid volume
summary(dispRity(dummy_space, metric = ellipse.volume))
```
> WARNING: in such dummy space, this gives the estimation of the ellipsoid volume, not the real ellipsoid volume! See the cautionary note in `?ellipse.volume`.

@@

```{r}
# Creating a 2D uniform space
example_space <- space.maker(300, 2, runif)

>>>>

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
1 | `disalignment` | The rejection of the centroid of a matrix from the major axis of another (typically an `"as.covar"` metric) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipsoid.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
1 | `projections.between` | `projections` metric applied between groups | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
1 | `roundness` | The integral of the ranked scaled eigenvalues of a variance-covariance matrix | `dispRity`
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

\begin{equation}
    ellipsoid.volume = \frac{\pi^{d/2}}{\Gamma(\frac{d}{2}+1)}\displaystyle\prod_{i=1}^{d} (\lambda_{i}^{0.5})
\end{equation}

@@

#### Volumes and surface metrics
The functions `ellipsoid.volume`, `convhull.surface`, `convhull.volume` and `n.ball.volume`  all measure the surface or the volume of the ordinated space occupied:

@@

```{r}
## Calculating the ellipsoid volume
summary(dispRity(dummy_space, metric = ellipsoid.volume))
```
> WARNING: in such dummy space, this gives the estimation of the ellipsoid volume, not the real ellipsoid volume! See the cautionary note in `?ellipsoid.volume`.

@@

```{r}
## Creating a 2D uniform space
example_space <- space.maker(300, 2, runif)

### 2023-12-06 inst/gitbook/06_guts.Rmd
updated book

#### `rescale.dispRity`

@@

## Scaling the same disparity values
head(summary(rescale.dispRity(disparity, scale = TRUE)))

## Scaling and centering:
head(summary(rescale.dispRity(disparity, scale = TRUE, center = TRUE)))

## Rescaling the value by dividing by a maximum value
head(summary(rescale.dispRity(disparity, max = 10)))
```

>>>>

#### `scale.dispRity`

@@

## Scaling the same disparity values
head(summary(scale.dispRity(disparity, scale = TRUE)))

## Scaling and centering:
head(summary(scale.dispRity(disparity, scale = TRUE, center = TRUE)))

## Rescaling the value by dividing by a maximum value
head(summary(scale.dispRity(disparity, max = 10)))
```
